---
title: "`ggplot` 1"
author: "Ashley Tseng"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
library(hexbin)
```
Note that `ggplot2` is automatically loaded once we load the tidyverse.
`ggridges` won't override any of the packages we have already loaded.

## Load dataset
```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
Inside the rNOAA package, there is a dataset that we want. This lets us interact with the NOAA servers and tell them what monitors we want data from. We will recode the IDs with `mutate`.


## Create a ggplot with minimum temp on the x-axis and max temp on the y axis:
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
We have only defined the x and y axes, but not the geometry (what we're plotting).

Alternate way of making this plot (via piping):
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```
Anytime you write code you should start with the data, and then do this, and then do this...
Note that this doesn't save the scatterplot.

Saving initial plots (prof. mostly doesn't do this):
```{r}
scatterplot = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

scatterplot
```


Adding color:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, alpha = 0.4))
```
alpha adds transparency

Why do `aes` positions matter?
First:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, alpha = 0.4)) +
  geom_smooth(se = FALSE)
```
First in ggplot we defined x and y.
Then we defined color in geom_point, so color matters ONLY inside geom_point but not to geom_smooth. So we see only one line for geom_point.
If we had defined color in ggplot, it would have applied to all geom_'s after that.

vs. 

Second:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```
Now we defined the color in ggplot and there's multiple smooth lines (one for each location).


## Facet!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(alpha = 0.4)) +
  geom_smooth(se = FALSE)  + 
  facet_grid(. ~ name)
```
Adding `facet_grid` results in three separate scatterplots per location.
This is fine, but not interesting. We've only plotted tmax and tmin. We would assume that as tmin increases, that tmax would increase too.

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE, alpha = 0.5) + 
  facet_grid(. ~ name)
```
We see that there's not much of a difference for Waikiki around the year.


# Some extra stuff
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE, alpha = 0.5)
```
You don't have to have ALL the geoms every time (the geoms that you pick are up to you).
Pick what's useful to you.

2d density
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```



## More kinds of plots!!

Univariate plots:
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```
Histogram: gives me some idea of distribution
Histogram has two types of aes preferences. One is the outline of the bars (color), one is fill (f)
Don't need to define a y-axis for histograms.

3 separate histograms:
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```



Density plots:
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```
Can make comparisons across locations more easily.
Prof's favorite of the histograms.



Box plots:
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```
For each distribution, we are getting an individual box plot per location.



Violin plots (not preferrable or particularly useful):
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin()

weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 4)
```
Violin plots could be useful to see if a particular group was bimodal.



Ridge plots:
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```
What do we see? We saw in geom_density that these densities were stacked on top of one another. Ridge plots show us the densities per location.



## Saving and embedding plots
```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```
DON'T use the built-in “Export” button! If you do, your figure is not reproducible – no one will know how your plot was exported. 
Instead, use ggsave().

