---
title: "`ggplot` 1"
author: "Ashley Tseng"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```
Note that `ggplot2` is automatically loaded once we load the tidyverse.
`ggridges` won't override any of the packages we have already loaded.

## Load dataset
```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
Inside the rNOAA package, there is a dataset that we want. This lets us interact with the NOAA servers and tell them what monitors we want data from. We will recode the IDs with `mutate`.


## Create a ggplot with minimum temp on the x-axis and max temp on the y axis:
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
We have only defined the x and y axes, but not the geometry (what we're plotting).

Alternate way of making this plot (via piping):
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```
Anytime you write code you should start with the data, and then do this, and then do this...
Note that this doesn't save the scatterplot.

Saving initial plots (prof. mostly doesn't do this):
```{r}
scatterplot = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

scatterplot
```


Adding color:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, alpha = 0.4))
```
alpha adds transparency

Why do `aes` positions matter?
First:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name, alpha = 0.4)) +
  geom_smooth(se = FALSE)
```
First in ggplot we defined x and y.
Then we defined color in geom_point, so color matters ONLY inside geom_point but not to geom_smooth. So we see only one line for geom_point.
If we had defined color in ggplot, it would have applied to all geom_'s after that.

vs. 

Second:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```
Now we defined the color in ggplot and there's multiple smooth lines (one for each location).


## Facet!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(alpha = 0.4)) +
  geom_smooth(se = FALSE)  + 
  facet_grid(. ~ name)
```
Adding `facet_grid` results in three separate scatterplots per location.
This is fine, but not interesting. We've only plotted tmax and tmin. We would assume that as tmin increases, that tmax would increase too.

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE, alpha = 0.5) + 
  facet_grid(. ~ name)
```
We see that there's not much of a difference for Waikiki around the year.


## Some extra stuff
```{r}

```


